---
title: "Doing Data Science"
author: "Thijs Leufkens"
date: "31-12-2016"
output: html_document
---

## all packages which are needed

```{r loadpackages, echo=FALSE}
## install.packages("data.table")
## install.packages("ggplot2")
## install.packages("tidyr")
```

This post will describe the EDA (Exploratory Data Analysis) excercise from the Doing Data Science book by Cathy O'Neil & Rachel Schutt.
The data represents a simulated day's worth of ads shown and clicks recorded on the New York Times home page in May 2012.

```{r loaddata}
data1 <- read.csv(url("http://stat.columbia.edu/~rachel/datasets/nyt1.csv"))
```

## What's in the dataset?

Let's check how many columns / rows etc. there are and plot the first to rows to see some of its content with our own eyes.

```{r firstlook, echo=TRUE}
dim(data1)
str(data1)
head(data1,2)
```

Ok, so age is logged as a 0 or 1 value. From the book we know that 0 = female and 1 = male, so that mystery is solved.

The first excercise in the book is to create a new variable age_group to categorize the users. A quick search on Google provided a topic on stackoverflow explaining a method to do so.
(http://stackoverflow.com/questions/12979456/r-code-to-categorize-age-into-group-bins-breaks)

```{r agegroups, echo=TRUE}
library(data.table)
agebreaks <- c(0,18,25,35,45,55,65,500)
agelabels= c("<18","18-24","25-34","35-44","45-54","55-64","65+")

## setDT(data1)[ , Age_Group:= cut(Age, breaks= agebreaks, right= FALSE, labels= agelabels)]
data1$Age_Group <- cut(data1$Age, breaks= agebreaks, right= FALSE, labels= agelabels)
```

Now that we have categorized users by age group, we can do some initial analysis to see if something stands out.
We are interested in the click-through-rate (CTR = clicks / impressions). The assumption is that if we have a click without an impression, the data is wrong.

```{r CTR, echo=TRUE}
data1$CTR <- data1$Clicks/data1$Impressions
```

Let's plot the number of impressions and CTR for each age group:

```{r plot1, echo=TRUE}
library("tidyr")
library("ggplot2")


ggplot(subset(data1, Impressions>0), aes(x=Impressions, fill=Age_Group))+geom_histogram(binwidth = 1)
ggplot(subset(data1, Impressions>0), aes(x=Clicks, fill=Age_Group))+geom_histogram(binwidth = 1)
ggplot(subset(data1, Impressions>0), aes(x=Age_Group, y = CTR, fill=Age_Group))+geom_boxplot()


```


```{r plot2, echo=TRUE}
subset(data1, Impressions>0) %>%
  gather(key, value, Impressions, Clicks) %>%
  ggplot(aes(x = Age_Group, y = value, colour=key)) + geom_bar(stat = "identity")

```

When we plot the number of clicks only, we notice something strange. 

```{r plot3, echo=TRUE}
subset(data1, Impressions>0) %>%
  gather(key, value, Clicks) %>%
  ggplot(aes(x = Age_Group, y = value, colour=key)) + geom_bar(stat = "identity")

```

The under 18 bucket seems to be an outlier, what's going on here?
Let's compared signed in vs not signed in:

```{r plot4, echo=TRUE}
subset(data1, Signed_In < 1) %>%
  gather(key, value, Clicks) %>%
  ggplot(aes(x = Age_Group, y = value, colour=key)) + geom_bar(stat = "identity")

subset(data1, Signed_In > 0) %>%
  gather(key, value, Clicks) %>%
  ggplot(aes(x = Age_Group, y = value, colour=key)) + geom_bar(stat = "identity")
```

Appartently the "not signed in" users are in the <18 bucket, let's explore a bit furhter

```{r Not_Signed_in, echo=TRUE}
anonymousdata1 <- subset(data1, Signed_In < 1)
min(anonymousdata1$Age)
max(anonymousdata1$Age)
min(anonymousdata1$Gender)
max(anonymousdata1$Gender)
min(anonymousdata1$Impressions)
max(anonymousdata1$Impressions)
min(anonymousdata1$Clicks)
max(anonymousdata1$Clicks)
```

As to be expected, no age and gender data exists for the anonymous users.
